Episodic memory: mental time travel or a quantum `memory wave' function?

### 2020-09-23 paper/main.tex
working on R2 comments

One practical way to distinguish content from context representations is to separate mental (or neural) representations that drift rapidly \citep[content;][]{PolyEtal05a, MannEtal12} or gradually \citep[context;][]{PolyEtal05a, MannEtal11, HowaEtal12, LohnEtal18, LongKaha18, FolkEtal18}.  An intuition for why contextual representation might be expected to evolve gradually is laid out by \cite{PolyKaha08}.  Essentially, slowly drifting thoughts may be used to ``index'' more rapidly drifting thoughts that are temporally proximal.  This notion, that our brain maintains parallel mental representations drifting at different time scales, has been well-characterized in a series of elegant fMRI and ECoG studies by Uri Hasson and colleagues~\citep{HassEtal08, LernEtal11, HoneEtal12a, AlyEtal18}.  Subsequent work by the same group has shown that this hierarchy of drifting mental representations plays a central role in how we remember continuous experiences and segment our experiences into discrete events~\citep{BaldEtal17}. This work mirrors the discovery of \textit{time cells} in the rodent hippocampus that respond preferentially to a specific time relative to a temporal reference (e.g., time elapsed since starting to run on an exercise wheel), whereby each cell appears to integrate incoming information at a different rate~\citep{PastEtal08, MacDEtal11} and the population activity serves to represent information at a range of timescales~\citep{MauEtal18}. Other work suggests that the lateral entorhinal cortex also plays a role in integrating information across a wide range of timescales~\citep{TsaoEtal18}, supporting precise temporal recall~\citep{MontEtal19}.  These ideas have been formalized in a family of theoretical models over the past several decades~\citep{HowaKaha02a, DianEtal07, SedeEtal08, PolyEtal09, ShanEtal09, ShanHowa10, ShanHowa12, HowaEtal14, Rang18}.  In turn, these models were inspired in part by earlier theoretical work suggesting that temporally proximal experiences become linked through their shared contextual features~\citep{Este55a,AtkiShif68}.  Other recent work indicates that different aspects of the reinstatement process itself might also occur at multiple timescales.  For example, \cite{LindEtal19} found that high-level semantic information was reinstated prior to low-level (e.g. perceptual) details in an associative memory task.  Collectively, these experimental and theoretical studies make two important contributions to our understanding of how we remember our past.  First, the studies suggest that our ongoing experiences are ``blurred out'' in time by neural integrators (where different brain structures or sub-structures integrate with different time constants).  Second, the studies suggest that members of this family of representations drifting at different rates become bound together such that when we retrieve memories about our past, we reactivate representations that carry information at a range of time scales.  These reactivations explain how our brains reactivate prior contexts when we remember the past.

>>>>

One practical way to distinguish content from context representations is to separate mental (or neural) representations that drift rapidly \citep[content;][]{PolyEtal05a, MannEtal12} or gradually \citep[context;][]{PolyEtal05a, MannEtal11, HowaEtal12, LohnEtal18, LongKaha18, FolkEtal18}.  An intuition for why contextual representation might be expected to evolve gradually is laid out by \cite{PolyKaha08}.  Essentially, slowly drifting thoughts may be used to ``index'' more rapidly drifting thoughts that are temporally proximal.  This notion, that our brain maintains parallel mental representations drifting at different time scales, has been well-characterized in a series of elegant fMRI and ECoG studies by Uri Hasson and colleagues~\citep{HassEtal08, LernEtal11, HoneEtal12a, AlyEtal18}.  Subsequent work by the same group has shown that this hierarchy of drifting mental representations plays a central role in how we remember continuous experiences and segment our experiences into discrete events~\citep{BaldEtal17}.  This body of work suggests that different brain regions reflect \textit{temporal receptive windows} that characterize the timescale(s) to which that region is maximally sensitive.  This research complements work on the \textit{temporal window of integration} in the speech and perception literatures~\citep[e.g.,][]{ViemWake91, GautEtal12, vanWEtal07}.  The temporal window of integration refers to the duration over which physically distinct stimuli are treated (by some brain system or network) as a single percept or an atomic unit.  This systems-level work mirrors the discovery of \textit{time cells} in the rodent hippocampus that respond preferentially to a specific time relative to a temporal reference (e.g., time elapsed since starting to run on an exercise wheel), whereby each cell appears to integrate incoming information at a different rate~\citep{PastEtal08, MacDEtal11} and the population activity serves to represent information at a range of timescales~\citep{MauEtal18}. Other work suggests that the lateral entorhinal cortex also plays a role in integrating information across a wide range of timescales~\citep{TsaoEtal18}, supporting precise temporal recall~\citep{MontEtal19}.  These ideas have been formalized in a family of theoretical models over the past several decades~\citep{HowaKaha02a, DianEtal07, SedeEtal08, PolyEtal09, ShanEtal09, ShanHowa10, ShanHowa12, HowaEtal14, Rang18}.  In turn, these models were inspired in part by earlier theoretical work suggesting that temporally proximal experiences become linked through their shared contextual features~\citep{Este55a,AtkiShif68}.  Other recent work indicates that different aspects of the reinstatement process itself might also occur at multiple timescales.  For example, \cite{LindEtal19} found that high-level semantic information was reinstated prior to low-level (e.g. perceptual) details in an associative memory task.  Collectively, these experimental and theoretical studies make two important contributions to our understanding of how we remember our past.  First, the studies suggest that our ongoing experiences are ``blurred out'' in time by neural integrators (where different brain structures or sub-structures integrate with different time constants).  Second, the studies suggest that representations that reflect different timescales become bound together such that when we retrieve memories about our past, we reactivate representations that carry information at a range of time scales.  These reactivations explain how our brains reactivate prior contexts when we remember the past.
